Equivalence Testing and the Second Generation P-Value

### 2018-08-24 manuscript.Rmd
feedback Marie

author_note: |
  All code associated with this article, including the reproducible manuscript, is available from https://github.com/Lakens/TOST_vs_SGPV.

@@

Second generation *p*-values (SGPV) were recently proposed to as a descriptive statistic that represents ‘the proportion of data-supported hypotheses that are also null hypotheses’ [@blume_second-generation_2018]. 
The researcher specifies an equivalence range around a null hypothesis of values that are considered practically equivalent to the null hypothesis. 
The SGPV measures the degree to which a set of data-supported parameter values falls within the interval null hypothesis. 
If the estimation interval falls completely within the equivalence range, the SGPV is 1. 
If the confidence interval falls completely outside of the equivalence range, the SGPV is 0. Otherwise the SGPV is a value between 0 and 1 that expresses the overlap of data-supported hypotheses and the equivalence range. 
When calculating the SGPV the set of data-supported parameter values can be represented by a confidence interval (CI), although other one could also choose to use credible intervals or Likelihood support intervals (SI). 
When a confidence interval is used, the SGPV and equivalence tests such as the Two One-Sided Tests (TOST) procedure [@lakens_equivalence_2017;@meyners_equivalence_2012;@schuirmann_comparison_1987] appear to have close ties, because both tests compare a confidence interval against an equivalence range.
Here, we aim to examine the similarities and differences between the TOST procedure and the SGPV.

@@

The second generation *p*-value (SGPV) is calculated as:
$$
  p _ { \delta } = \frac { \left| I \cap H _ { 0 } \right| } { | I | } \times \max \left\{ \frac { | I | } { 2 \left| H _ { 0 } \right| } , 1 \right\}
$$
where I is the interval based on the data (e.g., a 95% confidence interval) and H~0~ is the equivalence range. The first term of this formula implies that the second generation *p*-value is the width of the confidence interval that overlaps with the equivalence range, divided by the total width of the confidence interval. The second term is a 'small sample correction' (which will be discussed later) that comes into play whenever the confidence interval is more than twice as wide as the equivalence range.

To examine the relation between the TOST *p*-value and the SGPV we can calculate both statistics across a range of observed effect sizes. In Figure \ref{fig:TOSTSGPV1} *p*-values are plotted for the TOST procedure and the SGPV. The statistics are calculated for hypothetical one-sample *t*-tests for observed means ranging from 140 to 150 (on the x-axis). The equivalence range is set to 145 ± 2 (i.e., an equivalence range from 143 to 147), the observed standard deviation is assumed to be 2, and the sample size is 30. For example, for the left-most point in Figure \ref{fig:TOSTSGPV1} the SGPV and the TOST *p*-value is calculated for a hypothetical study with a sample size of 30, an observed standard deviation of 2, and an observed mean of 140, where the *p*-value for the equivalence test is 1, and the SGPV is 0. Our conclusions about the relationship between TOST *p*-values and SGPV in this article are not dependent upon any specific example, as readers can explore for themselves in an online Shiny app: http://shiny.ieis.tue.nl/TOST_vs_SGPV/.

```{r, include=FALSE}
step = 0.01

@@

  res <- TOSTone.raw(m = m, 
                     mu = mu,
                     sd = sd, 
                     n = n, 
                     low_eqbound = low_eqbound, 
                     high_eqbound = high_eqbound, 
                     alpha = alpha,
                     plot = FALSE,
                     verbose = FALSE
  )
  t <- (m - mu)/(sd/sqrt(n))
  t_list[count] <- t
  sgpv_list[count] <- p_delta(mu+res$LL_CI_TTEST, mu+res$UL_CI_TTEST, mu+low_eqbound, mu+high_eqbound)
  p_tost_list[count] <- max(res$TOST_p1, res$TOST_p2)
  p_list[count] <- 2 * pt(-abs(t), df = n-1)
}
```

```{r, TOSTSGPV1, fig.cap="Comparison of *p*-values from TOST (black line) and SGPV (dotted grey line) across a range of observed sample means (x-axis) tested against a mean of 145 in a one-sample *t*-test with a sample size of 30 and a standard deviation of 2."}
plot(NA, 
     ylim = c(0, 1), 
     xlim = c(0, 1001),
     yaxt = "n",
     xaxt = "n",
     ylab = "P-value for TOST and SGPV",
     xlab = "Observed Mean")
axis(1, at = seq(0,1000,100), labels = seq(140,150,1), las = 1)
axis(2, at = seq(0,1,0.1), labels = seq(0,1,0.1), las = 1)
lines(sgpv_list, type="l", col = "darkgrey", lwd = 3, lty = 3)
lines(p_tost_list, lwd = 3)
```

@@

tost_res2 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha,
                   plot = FALSE,
                   verbose = FALSE
)
SGPV2 <- TOST_to_SGPV(tost_res = tost_res2)
```

When the observed sample mean is 145, the sample size is 30, and the standard deviation is 2, and we are testing against equivalence bounds of 143 and 147 using the TOST procedure for a one-sample *t*-test, the equivalence test is significant, *t*(`r tost_res1$TOST_df`) = `r round(tost_res1$TOST_t1, digits = 2)`, *p* < .001. Because the 95% CI falls completely within the equivalence bounds, the SGPV is `r SGPV1` (see Figure \ref{fig:TOSTSGPV1}).

On the other hand, when the observed mean is 140, the equivalence test is not significant (the observed mean is far outside the equivalence range of 143 to 147), *t*(`r tost_res2$TOST_df`) = `r round(tost_res2$TOST_t1, digits = 2)`, *p* = `r round(tost_res2$TOST_p1, digits =  4)` (or more accurately, *p* > .999 as *p*-values are bounded between 0 and 1). Because the 95% CI falls completely outside the equivalence bounds, the SGPV is `r SGPV2` (see Figure  \ref{fig:TOSTSGPV1}).

@@

As the observed mean in a one-sample *t*-test lies closer to the test value (in Figure 4, from situation A to D, the mean gets closer to the test value by 0.1) the difference in the overlap changes uniformly.
```{r, include=FALSE}
m <- 146
mu <- 144.5
sd <- 500
n <- 1000000
low_eqbound = -2 
high_eqbound = 2 
alpha = 0.05

@@

par(new=TRUE)
curve(dnorm(x, 1.2, 0.5), from=-0.3, to=2.7, ylim = c(-2, 8), xlim = c(-3, 3), xlab = "", ylab = "", axes = FALSE)
par(new=TRUE)
curve(dnorm(x, 1.3, 0.5), from=-0.2, to=2.8, ylim = c(-4, 6), xlim = c(-3, 3), xlab = "", ylab = "", axes = FALSE)
par(new=TRUE)
curve(dnorm(x, 1.4, 0.5), from=-0.1, to=2.9, ylim = c(-6, 4), xlim = c(-3, 3), xlab = "", ylab = "", axes = FALSE)
par(new=TRUE)
curve(dnorm(x, 1.5, 0.5), from=0, to=3, ylim = c(-8, 2), xlim = c(-3, 3), xlab = "", ylab = "", axes = FALSE)
```

```{r, include=FALSE}
#For the plot above I wanted 95% CI using the TOST functions, so I adjusted the alpha level. But I need to use alpha = 0.05 to calculate the SGPV - so here I quickly recalculate the SGPV for in the text
m <- 146
mu <- 144.5
sd <- 500
n <- 1000000
low_eqbound = -2 
high_eqbound = 2 
alpha = 0.05

@@

tost_res4 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha,
                   plot = FALSE,
                   verbose = FALSE
)
SGPV4 <- TOST_to_SGPV(tost_res = tost_res4)
```
For example, the SGPV from A to D is `r round(SGPV1, digits = 2)`, `r round(SGPV2, digits = 2)`, `r round(SGPV3, digits = 2)`, and `r round(SGPV4, digits = 2)`. The difference in the percentage of overlap between A and B (`r round(SGPV1-SGPV2, digits = 2)`) is identical to the difference in the percentage of overlap between C and D as the mean gets 0.1 closer to the test value (`r round(SGPV3-SGPV4, digits = 2)`).
```{r , echo=FALSE}
tail1 <- 1-pnorm(2, 1.5, 0.5)
tail2 <- 1-pnorm(2, 1.4, 0.5)
tail3 <- 1-pnorm(2, 1.3, 0.5)
tail4 <- 1-pnorm(2, 1.2, 0.5)
```

As we move the observed mean closer to the test value in steps of 0.1 across A to D the *p*-value calculated for normally distributed data is not uniformly distributed. The probability of observing data more extreme than the upper bound of 2 is (from A to D) `r round(tail1, digits = 3)`, `r round(tail2, digits = 3)`, `r round(tail3, digits = 3)`, and `r round(tail4, digits = 3)`. As we can see, the difference between A and B (`r round(tail1-tail2, digits = 3)`) is not the same as the difference between C and D (`r round(tail3-tail4, digits = 3)`). Indeed, the difference in *p*-values is the largest as you start at *p* = 0.5 (when the observed mean falls on the test value), which is why the line in Figure \ref{fig:TOSTSGPV1} is the steepest at *p* = 0.5. Note that where the SGPV reaches 1 or 0, *p*-values closely approximate 0 and 1, but never reach these values.

@@

We can examine this situation by calculating the SGPV and performing the TOST for a situation where sample sizes are small and the equivalence range is narrow, such that the CI is more than twice as large as the equivalence range. 
```{r, include=FALSE}
step = 0.01

@@

The *p*-value from the TOST procedure and the SGPV are also unrelated when the CI is wider than the equivalence range (so the precision is low) and overlaps with the upper and lower equivalence bound, but the CI is *not* twice as wide as the equivalence range. In the example below, we see that the CI is only 1.79 times as wide as the equivalence bounds, but the CI overlaps with the lower and upper equivalence bounds. This means the SGPV is not set to 0.5, but it is constant across a range of observed means.
```{r, include=FALSE}
m <- 0.2
n <- 10
sd <- 1
d.lo <- -0.4
d.hi <- 0.4

@@

If we once more plot the two statistics against each other to see where they are unrelated (indicated by straight lines in the curve), we see the SGPV is 0.56 for a range of observed means where the *p*-value from the equivalence test still varies.

@@

The effect of asymmetric confidence intervals is most easily noticable at smaller sample sizes, therefore in Figure \ref{fig:TOSTSGPV11} below we plot the *p*-values from equivalence tests and the SGPV (again plotted as 1-SGPV for ease of comparison) for correlations. The sample size is 30 pairs of observations, and the lower and upper equivalence bounds are set to -0.45 and 0.45, with an alpha of 0.05. As the observed correlation in the sample moves from -1 to 0 the *p*-value from the equivalence test becomes smaller, as does 1-SGPV. The pattern is quite similar to that in Figure \ref{fig:TOSTSGPV2}. The *p*-value for the TOST procedure and 1-SGPV are still identical when *p*-values are 0.975 and 0.025 (indicated by the upper and lower horizontal dotted lines). There are two important differences, however. First of all, the SGPV is no longer a straight line, but a curve, due to the asymmetry in the 95% CI. Second, and most importantly, the *p*-value for the equivalence test and the SGPV do no longer overlap at *p* = 0.5.

@@

The reason that the equivalence test and SGPV no longer overlap is also because of asymmetric confidence intervals. If the observed correlation falls exactly on the equivalence bound the *p*-value for the equivalence test indicates that the probability of observing the observed or more extreme data, assuming the equivalence bound is the true effect size, is 50%. In other words, if the true effect size is the same as the equivalence bound, it is equally likely to find an effect more extreme than the equivalence bound, as it is to observe an effect that is less extreme than the equivalence bound. However, as can be seen in Figure \ref{fig:TOSTSGPV12}, the two second generation *p*-values associated with the observed correlations at r = -0.45 and r = 0.45 are `r round(sgpv,3)`. Because the confidence intervals are asymmetric around the observed effect size of 0.45 (ranging from `r round(res$LL_CI_TTEST,2)` to `r round(res$UL_CI_TTEST,2)`) according to @blume_second-generation_2018 `r round(sgpv,3)*100`% of the data-supported hypotheses are null hypotheses, and therefore `r round(sgpv,3)*100`% of the data-supported hypotheses are compatible with the null premise.

@@

When introducing a new statistical method, it is important to compare it to existing approaches and specify it's relative strengths and weaknesses. First of all, the SGPV is a descriptive statistic (unlike the *p*-value that is calculated for an equivalence test, which is an inferential statistic) that provides the proportion of overlap of the confidence interval and the equivalence range.

Even though a SGPV of 1 or 0 has a clear interpretation (we can reject effects outside or inside the equivalence range), intermediate values are not as easy to interpret (especially for effects that have asymmetric confidence intervals). In one sense, they are what they are (the proportion of overlap), but in any specific cases it can be unclear what this number tells us about the data we have collected. This is not too problematic, since the main use of the SGPV (e.g., in all examples provided by Blume and colleagues) is to examine whether the SGPV is 0, 1, or inconclusive. This interpretation of a SGPV as allowing researchers to reject the null, reject the presence of a meaningful effect, or remaining inconclusive is very similar to the Neyman-Pearson interpretation of combining a null-hypothesis test and an equivalence test (@lakens_equivalence_2018). The difference is that where a SGPV of 1 can be interpreted as *p* < .025, equivalence tests provide exact *p*-values, and they continue to differentiate between for example p = 0.048 and p = 0.002. Whether this is desirable depends on the perspective that is used. From a Neyman-Pearson perspective on statistical inferences the main conclusion is based on whether or not $p < \alpha$, and thus an equivalence test and SGPV can be performed by simply checking whether the confidence interval falls within the equivalence range, just as a null-hypothesis test can be performed by checking whether the confidence interval contains zero or not. At the same time, it is recommended to report exact *p*-values [@american_psychological_association_publication_2010], and exact *p*-values might provide information of interest to readers about how precisely how surprising the data is under the null model. Equivalence tests combined with null-hypothesis significance tests also allow researchers to conclude an effect is significant *and* equivalent (i.e., statistically different from zero, but also too small to be considered meaningful). Thus, the SGPV is used to classify results into one of three possible categories (with the data falling inside or outside the equivalence range, or being inconclusive), while equivalence tests combined with null-hypothesis tests classify results into four possible categories.

@@

We believe that our explanation of the similarities between the TOST procedure and the SGPV provides context to interpret the contribution of second generation *p*-values to the statistical toolbox. The novelty of te SGPV lies in its use as a descriptive statistic, but this use can be limited when confidence intervals are asymmetrical or wider than the equivalence range. There are strong similarities with *p*-values from the TOST procedure, and in all situations where the statistics yield different results, the behavior of the *p*-value from the TOST procedure is more consistent and easier to interpret. We hope this overview of the relationship between the SGPV and equivalence tests will help researchers to make an informed decision about which statistical approach provides the best answer to their question. Our comparisons shows that when proposing alternatives to null-hypothesis tests, it is important to compare new proposals to already existing procedures. We believe equivalence tests achieve the goals of the second generation *p*-value while allowing users to more easily control error rates, and while yielding more consistent statistical outcomes.

>>>>

author_note: |
  All code associated with this article, including the reproducible manuscript, is available from https://github.com/Lakens/TOST_vs_SGPV. This work was supported by the Netherlands Organization for Scientific Research (NWO) VIDI grant 452-17-013.

@@

Second generation *p*-values (SGPV) were recently proposed to as a descriptive statistic that represents ‘the proportion of data-supported hypotheses that are also null hypotheses’ [@blume_second-generation_2018]. 
The researcher specifies an equivalence range around a null hypothesis of values that are considered practically equivalent to the null hypothesis. 
The SGPV measures the degree to which a set of data-supported parameter values falls within the interval null hypothesis. 
If the estimation interval falls completely within the equivalence range, the SGPV is 1. 
If the confidence interval falls completely outside of the equivalence range, the SGPV is 0. Otherwise the SGPV is a value between 0 and 1 that expresses the overlap of data-supported hypotheses and the equivalence range. 
When calculating the SGPV the set of data-supported parameter values can be represented by a confidence interval (CI), although one could also choose to use credible intervals or Likelihood support intervals (SI). 
When a confidence interval is used, the SGPV and equivalence tests such as the Two One-Sided Tests (TOST) procedure [@lakens_equivalence_2017;@meyners_equivalence_2012;@quertemont_how_2011;@schuirmann_comparison_1987] appear to have close ties, because both tests compare a confidence interval against an equivalence range.
Here, we aim to examine the similarities and differences between the TOST procedure and the SGPV.

@@

The second generation *p*-value (SGPV) is calculated as:
$$
  p _ { \delta } = \frac { \left| I \cap H _ { 0 } \right| } { | I | } \times \max \left\{ \frac { | I | } { 2 \left| H _ { 0 } \right| } , 1 \right\}
$$
where I is the interval based on the data (e.g., a 95% confidence interval) and H~0~ is the equivalence range. The first term of this formula implies that the second generation *p*-value is the width of the confidence interval that overlaps with the equivalence range, divided by the total width of the confidence interval. The second term is a 'small sample correction' (which will be discussed later) that comes into play whenever the confidence interval is more than twice as wide as the equivalence range.
To examine the relation between the TOST *p*-value and the SGPV we can calculate both statistics across a range of observed effect sizes. In Figure \ref{fig:TOSTSGPV1} *p*-values are plotted for the TOST procedure and the SGPV. The statistics are calculated for hypothetical one-sample *t*-tests for observed means ranging from 140 to 150 (on the x-axis). The equivalence range is set to 145 ± 2 (i.e., an equivalence range from 143 to 147), the observed standard deviation is assumed to be 2, and the sample size is 30. For example, for the left-most point in Figure \ref{fig:TOSTSGPV1} the SGPV and the TOST *p*-value is calculated for a hypothetical study with a sample size of 30, an observed standard deviation of 2, and an observed mean of 140, where the *p*-value for the equivalence test is 1, and the SGPV is 0. Our conclusions about the relationship between TOST *p*-values and SGPV in this article are not dependent upon any specific example, as readers can explore for themselves in an online Shiny app: http://shiny.ieis.tue.nl/TOST_vs_SGPV/.
```{r, include=FALSE}
step = 0.01

@@

  res <- TOSTone.raw(m = m, 
                     mu = mu,
                     sd = sd, 
                     n = n, 
                     low_eqbound = low_eqbound, 
                     high_eqbound = high_eqbound, 
                     alpha = alpha,
                     plot = FALSE,
                     verbose = FALSE
  )
  t <- (m - mu)/(sd/sqrt(n))
  t_list[count] <- t
  sgpv_list[count] <- p_delta(mu+res$LL_CI_TTEST, mu+res$UL_CI_TTEST, mu+low_eqbound, mu+high_eqbound)
  p_tost_list[count] <- max(res$TOST_p1, res$TOST_p2)
  p_list[count] <- 2 * pt(-abs(t), df = n-1)
}
```
```{r, TOSTSGPV1, fig.cap="Comparison of *p*-values from TOST (black line) and SGPV (dotted grey line) across a range of observed sample means (x-axis) tested against a mean of 145 in a one-sample *t*-test with a sample size of 30 and a standard deviation of 2."}
plot(NA, 
     ylim = c(0, 1), 
     xlim = c(0, 1001),
     yaxt = "n",
     xaxt = "n",
     ylab = "P-value for TOST and SGPV",
     xlab = "Observed Mean")
axis(1, at = seq(0,1000,100), labels = seq(140,150,1), las = 1)
axis(2, at = seq(0,1,0.1), labels = seq(0,1,0.1), las = 1)
lines(sgpv_list, type="l", col = "darkgrey", lwd = 3, lty = 3)
lines(p_tost_list, lwd = 3)
```

@@

tost_res2 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha,
                   plot = FALSE,
                   verbose = FALSE
)
SGPV2 <- TOST_to_SGPV(tost_res = tost_res2)
```
When the observed sample mean is 145, the sample size is 30, and the standard deviation is 2, and we are testing against equivalence bounds of 143 and 147 using the TOST procedure for a one-sample *t*-test, the equivalence test is significant, *t*(`r tost_res1$TOST_df`) = `r round(tost_res1$TOST_t1, digits = 2)`, *p* < .001. Because the 95% CI falls completely within the equivalence bounds, the SGPV is `r SGPV1` (see Figure \ref{fig:TOSTSGPV1}). 
On the other hand, when the observed mean is 140, the equivalence test is not significant (the observed mean is far outside the equivalence range of 143 to 147), *t*(`r tost_res2$TOST_df`) = `r round(tost_res2$TOST_t1, digits = 2)`, *p* = `r round(tost_res2$TOST_p1, digits =  4)` (or more accurately, *p* > .999 as *p*-values are bounded between 0 and 1). Because the 95% CI falls completely outside the equivalence bounds, the SGPV is `r SGPV2` (see Figure  \ref{fig:TOSTSGPV1}).

@@

```{r, include=FALSE}
m <- 146
mu <- 144.5
sd <- 500
n <- 1000000
low_eqbound = -2 
high_eqbound = 2 
alpha = 0.05

@@

par(new=TRUE)
curve(dnorm(x, 1.2, 0.5), from=-0.3, to=2.7, ylim = c(-2, 8), xlim = c(-3, 3), xlab = "", ylab = "", axes = FALSE)
par(new=TRUE)
curve(dnorm(x, 1.3, 0.5), from=-0.2, to=2.8, ylim = c(-4, 6), xlim = c(-3, 3), xlab = "", ylab = "", axes = FALSE)
par(new=TRUE)
curve(dnorm(x, 1.4, 0.5), from=-0.1, to=2.9, ylim = c(-6, 4), xlim = c(-3, 3), xlab = "", ylab = "", axes = FALSE)
par(new=TRUE)
curve(dnorm(x, 1.5, 0.5), from=0, to=3, ylim = c(-8, 2), xlim = c(-3, 3), xlab = "", ylab = "", axes = FALSE)
```
```{r, include=FALSE}
#For the plot above I wanted 95% CI using the TOST functions, so I adjusted the alpha level. But I need to use alpha = 0.05 to calculate the SGPV - so here I quickly recalculate the SGPV for in the text
m <- 146
mu <- 144.5
sd <- 500
n <- 1000000
low_eqbound = -2 
high_eqbound = 2 
alpha = 0.05

@@

tost_res4 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha,
                   plot = FALSE,
                   verbose = FALSE
)
SGPV4 <- TOST_to_SGPV(tost_res = tost_res4)
```
```{r , echo=FALSE}
tail1 <- 1-pnorm(2, 1.5, 0.5)
tail2 <- 1-pnorm(2, 1.4, 0.5)
tail3 <- 1-pnorm(2, 1.3, 0.5)
tail4 <- 1-pnorm(2, 1.2, 0.5)
```
For example, the SGPV from A to D is `r round(SGPV1, digits = 2)`, `r round(SGPV2, digits = 2)`, `r round(SGPV3, digits = 2)`, and `r round(SGPV4, digits = 2)`. The difference in the percentage of overlap between A and B (`r round(SGPV1-SGPV2, digits = 2)`) is identical to the difference in the percentage of overlap between C and D as the mean gets 0.1 closer to the test value (`r round(SGPV3-SGPV4, digits = 2)`).
As the observed mean in a one-sample *t*-test lies closer to the test value (in Figure \ref{fig:TOSTSGPV4}), from situation A to D, the mean gets closer to the test value by 0.1) the difference in the overlap changes uniformly. 
As we move the observed mean closer to the test value in steps of 0.1 across A to D the *p*-value calculated for normally distributed data is not uniformly distributed. The probability of observing data more extreme than the upper bound of 2 is (from A to D) `r round(tail1, digits = 3)`, `r round(tail2, digits = 3)`, `r round(tail3, digits = 3)`, and `r round(tail4, digits = 3)`. As we can see, the difference between A and B (`r round(tail1-tail2, digits = 3)`) is not the same as the difference between C and D (`r round(tail3-tail4, digits = 3)`). Indeed, the difference in *p*-values is the largest as you start at *p* = 0.5 (when the observed mean falls on the test value), which is why the line in Figure \ref{fig:TOSTSGPV1} is the steepest at *p* = 0.5. Note that where the SGPV reaches 1 or 0, *p*-values closely approximate 0 and 1, but never reach these values.

@@

We can examine this situation by calculating the SGPV and performing the TOST for a situation where sample sizes are small and the equivalence range is narrow, such that the CI is more than twice as large as the equivalence range (see Figure \ref{fig:TOSTSGPV6}). 
```{r, include=FALSE}
step = 0.01

@@

The *p*-value from the TOST procedure and the SGPV are also unrelated when the CI is wider than the equivalence range (so the precision is low) and overlaps with the upper and lower equivalence bound, but the CI is *not* twice as wide as the equivalence range. In the example below, we see that the CI is only 1.79 times as wide as the equivalence bounds, but the CI overlaps with the lower and upper equivalence bounds (Figure \ref{fig:TOSTSGPV8}). This means the SGPV is not set to 0.5, but it is constant across a range of observed means.
```{r, include=FALSE}
m <- 0.2
n <- 10
sd <- 1
d.lo <- -0.4
d.hi <- 0.4

@@

If we once more plot the two statistics against each other to see where they are unrelated (indicated by straight lines in the curve), we see the SGPV is 0.56 for a range of observed means where the *p*-value from the equivalence test still varies (Figure \ref{fig:TOSTSGPV10}).

@@

The effect of asymmetric confidence intervals is most noticable at smaller sample sizes. In Figure \ref{fig:TOSTSGPV11} we plot the *p*-values from equivalence tests and the SGPV (again plotted as 1-SGPV for ease of comparison) for correlations. The sample size is 30 pairs of observations, and the lower and upper equivalence bounds are set to -0.45 and 0.45, with an alpha of 0.05. As the observed correlation in the sample moves from -1 to 0 the *p*-value from the equivalence test becomes smaller, as does 1-SGPV. The pattern is quite similar to that in Figure \ref{fig:TOSTSGPV2}. The *p*-value for the TOST procedure and 1-SGPV are still identical when *p*-values are 0.975 and 0.025 (indicated by the upper and lower horizontal dotted lines). There are two important differences, however. First of all, the SGPV is no longer a straight line, but a curve, due to the asymmetry in the 95% CI. Second, and most importantly, the *p*-value for the equivalence test and the SGPV do no longer overlap at *p* = 0.5.

@@

The reason that the equivalence test and SGPV no longer overlap is also because of asymmetric confidence intervals. If the observed correlation falls exactly on the equivalence bound the *p*-value for the equivalence test indicates that the probability of observing the observed or more extreme data, assuming the equivalence bound is the true effect size, is 50%. In other words, if the true effect size is the same as the equivalence bound, it is equally likely to find an effect more extreme than the equivalence bound, as it is to observe an effect that is less extreme than the equivalence bound. However, as can be seen in Figure \ref{fig:TOSTSGPV12}, the two second generation *p*-values associated with the observed correlations at r = -0.45 and r = 0.45 are larger than 50%. Because the confidence intervals are asymmetric around the observed effect size of 0.45 (ranging from `r round(res$LL_CI_TTEST,2)` to `r round(res$UL_CI_TTEST,2)`) according to @blume_second-generation_2018 `r round(sgpv*100,3)`% of the data-supported hypotheses are null hypotheses, and therefore `r round(sgpv,3)*100`% of the data-supported hypotheses are compatible with the null premise.

@@

When introducing a new statistical method, it is important to compare it to existing approaches and specify it's relative strengths and weaknesses. First of all, the SGPV is a descriptive statistic (unlike the *p*-value that is calculated for an equivalence test, which is an inferential statistic) that provides the proportion of overlap of the confidence interval and the equivalence range. Even though a SGPV of 1 or 0 has a clear interpretation (we can reject effects outside or inside the equivalence range), intermediate values are not as easy to interpret (especially for effects that have asymmetric confidence intervals). In one sense, they are what they are (the proportion of overlap), but in any specific cases it can be unclear what this number tells us about the data we have collected. This is not too problematic, since the main use of the SGPV (e.g., in all examples provided by Blume and colleagues) is to examine whether the SGPV is 0, 1, or inconclusive. This interpretation of a SGPV as allowing researchers to reject the null, reject the presence of a meaningful effect, or remaining inconclusive is very similar to the Neyman-Pearson interpretation of combining a null-hypothesis test and an equivalence test [@lakens_equivalence_2018]. The difference is that where a SGPV of 1 can be interpreted as *p* < .025, equivalence tests provide exact *p*-values, and they continue to differentiate between for example *p* = 0.048 and *p* = 0.002. Whether this is desirable depends on the perspective that is used. From a Neyman-Pearson perspective on statistical inferences the main conclusion is based on whether or not $p < \alpha$, and thus an equivalence test and SGPV can be performed by simply checking whether the confidence interval falls within the equivalence range, just as a null-hypothesis test can be performed by checking whether the confidence interval contains zero or not. At the same time, it is recommended to report exact *p*-values [@american_psychological_association_publication_2010], and exact *p*-values might provide information of interest to readers about how precisely how surprising the data is under the null model. Equivalence tests combined with null-hypothesis significance tests also allow researchers to conclude an effect is significant *and* equivalent (i.e., statistically different from zero, but also too small to be considered meaningful). Thus, the SGPV is used to classify results into one of three possible categories (with the data falling inside or outside the equivalence range, or being inconclusive), while equivalence tests combined with null-hypothesis tests classify results into four possible categories.

@@

We believe that our explanation of the similarities between the TOST procedure and the SGPV provides context to interpret the contribution of second generation *p*-values to the statistical toolbox. The novelty of the SGPV lies in its use as a descriptive statistic, but this use can be limited when confidence intervals are asymmetrical or wider than the equivalence range. There are strong similarities with *p*-values from the TOST procedure, and in all situations where the statistics yield different results, the behavior of the *p*-value from the TOST procedure is more consistent and easier to interpret. We hope this overview of the relationship between the SGPV and equivalence tests will help researchers to make an informed decision about which statistical approach provides the best answer to their question. Our comparisons shows that when proposing alternatives to null-hypothesis tests, it is important to compare new proposals to already existing procedures. We believe equivalence tests achieve the goals of the second generation *p*-value while allowing users to more easily control error rates, and while yielding more consistent statistical outcomes.
